The Sub-Millimetre User Reduction Facility: an iterative map-maker for SCUBA-2

### 2011-10-21 smurf_paper.tex
edits to intro

\author[Edward~L.~Chapin~et~al.]{
  \parbox[t]{\textwidth}{
    Edward~L.~Chapin$^{1}$\thanks{E-mail:~echapin@phas.ubc.ca},
    David~S.~Berry$^{2}$,
    Andrew~G.~Gibb$^{1}$,
    Tim~Jenness$^{2}$,
    Douglas~Scott$^{1}$
  }
  \\
  \\
  $^{1}$Dept. of Physics \& Astronomy, University of British Columbia,
  6224 Agricultural Road, Vancouver, B.C. V6T 1Z1, Canada\\
  $^{2}$JointAstronomy Centre, 660 N. A`oh\={o}k\={u} Place, University
  Park, Hilo, Hawaii 96720, USA}

@@

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{holland2006} is a new instrument for the 15-m
James Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
can simultaneously image the sky in two broad bands centered over 450
and 850\,\micron. Ultimately, the instrument will also be equipped
with a Fourier transform spectrometer \citep[FTS-2,][]{gom2010}, and a polarimeter
\cite[POL-2,][]{bastien2005}. This paper describes the Submillimetre User Reduction Facility,
SMURF, a software package for reducing the imaging data, with an
emphasis on data taken during the SCUBA-2 Shared-Risks Observing period
(S2SRO) which took place from 2010 February 11 to 2010 March 25. The reduction
of FTS-2 and POL-2 data will be described at a later date once these
additional instruments have been commissioned.

Over the last twenty years, the submillimetre band (defined here to be
200--1000\,\micron) has revolutionized several important areas of
astrophysics: helping to characterize the early stages of
star-formation by identifying the dense, cold regions in molecular
clouds where stars may eventually form; discovering through blind
surveys the locations and surface density of a class of massive
star-forming galaxies in the early Universe, now referred to as
submillimetre galaxies, or SMGs; and finding debris disks around
nearby stars, the early stages of planet formation.

Styles of map-making: maximum likelihood
\citep[e.g.,][]{patanchon2008}; iterative -- see \citet{johnstone2000}
implementation of \citet{wright1996} to remove chop,
\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA analysis
\citep[e.g.][]{laurent2005,scott2008,aguirre2010}.

Since \scuba\ will ultimately have nearly 10,000 working bolometers,
considerably larger than any other existing ground, or space-spaced
bolometer cameras, SMURF has been designed to use a less accurate,
though faster and less memory-intensive iterative approach that
attempts to model and remove most of the correlated signal components,
and then regrid the residuals signals. This method is similar in
spirit to the Comprehensive Reduction Utility for SHARC-2
\citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
Bolocam Galactic Plane Survey \citep{aguirre2010}.

@@

Before describing the data-reduction algorithm in detail, we summarize
the basic strategy used to make measurements with SCUBA-2.

The first generation of submm bolometer cameras used NTD germanium
thermistors (e.g., SCUBA, MAMBO, BOLOCAM, AzTEC, LABOCA...).
Generally speaking, these devices are operated using a current bias;
changing thermal conditions alter their impedence, resulting in a
changing Voltage that can be measured. The bolometers are cooled so
that they are {\em background limited} -- the dominant noise term is
the optical load from sources outside of the cryostat (emissivity from
telescope, atmospheric fluctuations, and astronomical sources).

In contrast, SCUBA-2 belongs to a new generation of cameras that uses
superconducting transition-edge sensors (TESs, like ACT, SPT,
MUSTANG). These devices are held at the critical temperature,
$T_\mathrm{C}$, below which the material becomes a superconductor, and
above which it becomes a normal conductor. Within this transition,
tiny temperature fluctuations can produce huge changes in
impedence. The devices are operated with a voltage bias, so that
thermal fluctuations produce changing currents, which in turn generate
changing local magnetic fields. Chains of superconducting quantum
interference devices (SQUIDs) are used to amplify these signals,
ultimately providing a larger changing current that may be
digitized. Like the NTD thermistor-based cameras, SCUBA-2 is cooled in
order to be background limited. However, a superior over-all
sensitivity on a per-bolometer basis is achieved through its improved
responsivity (similar to an improved quantum efficiency for a CCD).

The load on the detectors can vary substantially, both from thermal
sources within and outside the SCUBA-2 cryostat. In order to operate
the camera across a broad dynamic range of conditions, it is therefore
necessary to thermally control the focal plane. This is achieved by
providing a fridge {\em base temperature}, $T_\mathrm{B}$ that is
well-below $T_\mathrm{C}$, and then using heaters that are connected
to each bolometer to place the array in to transition.

The focal plane is divided into four quadrants, each filled with 32
column $\times$ 40 row sub-arrays labelled s4a--s4d at 450\,\micron,
and s8a--s8d at 850\,\micron. To avoid independent bias and readout
wires to each bolometer, each array is multiplexed by column. There is
an additional row of dark SQUIDs with no TESs or thermal absorber used
to monitor non-thermal noise sources.

>>>>

\author[Edward~L.~Chapin~et~al.]{
  \parbox[t]{\textwidth}{
    Edward~L.~Chapin$^{1,2}$\thanks{E-mail:~echapin@phas.ubc.ca},
    David~S.~Berry$^{2}$,
    Andrew~G.~Gibb$^{1}$,
    Tim~Jenness$^{2}$,
    Douglas~Scott$^{1}$
  }
  \\
  \\
  $^{1}$Dept. of Physics \& Astronomy, University of British Columbia,
  6224 Agricultural Road, Vancouver, B.C. V6T 1Z1, Canada\\
  $^{2}$JointAstronomy Centre, 660 N. A`oh\={o}k\={u} Place, University
  Park, Hilo, Hawaii 96720, USA}

@@

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{holland2006} is a new instrument for the 15-m James
Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
simultaneously images the sky in two broad bands centered over 450 and
850\,\micron. The focal planes at each wavelength are populated with 4
rectangular subarrays, consisting of $40 \times 32$ bolometers each,
and subtend a nearly contiguous $4' \times 4'$ footprint on the sky.
Ultimately, the instrument will also be equipped with a Fourier
transform spectrometer \citep[FTS-2,][]{gom2010}, and a polarimeter
\cite[POL-2,][]{bastien2005}. This paper describes the Submillimetre
User Reduction Facility, SMURF, a software package for reducing the
imaging data. The reduction of FTS-2 and POL-2 data will be described
at a later date once these additional instruments have been
commissioned. The details of the instrument design, performance, and
its calibration are given in two companion papers: \citet{holland2011}
and \citet{dempsey2011}.

Over the last twenty years, observations in the submillimetre band
(defined here to be 200--1000\,\micron) have helped revolutionized
several important areas of astrophysics including: the
characterization of early stages of star-formation by identifying the
dense, cold regions in molecular clouds where stars may eventually
form; discovering through blind surveys a class of massive dusty
star-forming galaxies in the early ($z>2$) Universe, now referred to
as submillimetre galaxies, or SMGs; and identifying debris disks
around nearby stars, the early stages of planet formation.  With
$\sim10000$ total detectors, SCUBA-2 is presently the largest
submillimetre camera in the world, and the fastest wide-area
ground-based submillimetre imager.

Most submillimetre imaging cameras to date have used bolometers to
maximize sensitivities: each pixel consists of an efficient thermal
absorber to collect photons that have passed through a broad-band
filter, and the signal is read-out using a thermometer. The
high-frequency broad-band noise in modern systems, such as SCUBA-2, is
limited by photon noise from the instrument and ambient
backgrounds. The low-frequency noise, however, is dominated by sources
such as the low-frequency drift in the readout electronics, and drifts
in the mean background levels (such as the atmosphere for ground-based
instruments). The process of making a map therefore amounts to
separating the component of the bolometer signals that is produced by
astronomical sources, from the non-white (and therefore correlated)
contaminating noise components. Fortunately each has a particular
profile: astronomical sources are typically fixed in a sky reference
frame (to a good approximation), whereas other noise sources are often
correlated in time, and usually from bolometer-to-bolometer. ...also
mention scan strategies...

There are numerous ways to attack this problem, most of them evolving
out of techniques that were developed to produce maps of the
large-scale anisitropies in the Cosmic Microwave Backgroubd.  Styles
of map-making: maximum likelihood \citep[e.g.,][]{patanchon2008};
iterative -- see \citet{johnstone2000} implementation of
\citet{wright1996} to remove chop, \citet{kovacs2008} for SHARC-2
etc.; de-correlation using PCA analysis
\citep[e.g.][]{laurent2005,scott2008,aguirre2010}.

\scuba\ is presently unique in terms of the shear volume of data that
it produces on a nightly basis, and the fact that it is observing a
wide variety of science targets for multiple users (unlike dedicated
CMB experiments, for example). For this reason, SMURF has been
designed to use a less accurate (primarily on large angular scales),
though faster and less memory-intensive iterative approach that
attempts to model and remove most of the correlated signal components,
and then regrid the residual signals. This method is similar in spirit
to the Comprehensive Reduction Utility for SHARC-2
\citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
Bolocam Galactic Plane Survey \citep{aguirre2010}. While most of the
analysis described here applies equally to all of the data taken
to-date with \scuba, we note that the SCUBA-2 Shared-Risks Observing
period (S2SRO) which took place from 2010 February 11 to 2010 March 25
used only a single subarray at each of the two observing wavelengths,
and had considerably different noise properties and poorer performance
than the fully-commissioned instrument in 2011.

@@

While the details of how \scuba\ works is described in
\citet{holland2011}, and its calibration in \citet{dempsey2011}, we
summarize the basic concepts relative to map-making here.

The \scuba\ bolometer temperatures are read-out using superconducting
transition-edge sensors (TESs), with critical temperatures
$T_\mathrm{C}$ (...typical value...). The range about $T_\mathrm{C}$
in which the transition from a superconductor to a normal material
occurs is extremely small (...typical value...?). If the devices are
biased to a temperature within this transition on large time scales,
then tiny temperature fluctuations on short time-scales will produce
significant changes in the impedence. The entire SCUBA-2 focal plane
is therefore held at a \emph{base temperature} somewhat below
$T_\mathrm{C}$. The TESs then have a small electrical voltage bias,
and individual pixel heaters (using a current bias to apply the
desired Ohmic heating) are used to place the detectors in the
transition. Only a small detector bias voltage is required since the
heaters compensate for most of the variations in the base temperature,
and other ambient loads (like the atmosphere).  Small temperature (and
therefore impedence) fluctuations produce a changing current through
the TESs. The resulting changing magnetic fields are then amplified
using chains of superconducting quantum interference devices (SQUIDs),
and the larger output current is finally digitized. Each detector has
its own SQUID for the first-stage of the amplification, but the
remaining stages occur within a common chain of SQUIDs for each column
of 40 detectors. Each of the 32 rows is then read-out in sequence that
occurs at a frequency of ...kHz?... This multiplexing scheme greatly
reduces the number of readout and bias lines to each detector, making
such large-format bolometer array designs feasible. It should also be
noted that there is an additional 33rd row of SQUID reasouds that are
not connected to TES. These ``dark SQUIDs'' are used to track
non-thermal noise sources that are common to each column's amplifier
chain.

The focal planes at each of 450 and 850\,\micron\ are populated with
four of these $40 \times 32$ subarrays (referred to as s4a--s4d, and
s8a--s8d respectively), and simultaneously image approximately the
same region of the sky by means of a beam-splitter.
